*Symbiodinium* communities in reef corals of St. John, U.S.V.I. (2012)

### 2016-09-01 analysis/STJ2012_analysis.Rmd
update otu taxonomy data import and filtering

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = normalizePath(".."))
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 3, digits = 2)
```

```{r setup_2, include=FALSE}
# Load all R scripts in "R/" directory
sapply(list.files(path="R", pattern="*.R", full.names=TRUE), source, .GlobalEnv)
# Load package libraries
library(phyloseq); library(vegan); library(multcompView); library(reshape2); library(igraph); library(stringr)
# Set colors for plotting clades
taxcolors <- matrix(c("#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462"), 
                    dimnames=list(c("CladeA", "CladeB", "CladeC", "CladeD", "CladeF", "CladeG")))
```

@@

# Data pre-processing
### Import dataset
The [phyloseq](https://joey711.github.io/phyloseq/) package is used here to combine the OTU table, taxonomic assignments, and sample metadata into a single R data object (class 'phyloseq') to facilitate downstream analyses.
```{r data_import, cache=T}
# Import OTU tables
OTU97 <- otu_table(read.table("data/97_otus.tsv", header=T, check.names=F, row.names=1,
                              skip=1, sep="\t", comment.char=""), taxa_are_rows=T)
OTU97bs <- otu_table(read.table("data/97_otus_bysample.tsv", header=T, check.names=F, row.names=1,
                              skip=1, sep="\t", comment.char=""), taxa_are_rows=T)
OTU97bsp <- otu_table(read.table("data/97_otus_byspecies.tsv", header=T, check.names=F, row.names=1,
                              skip=1, sep="\t", comment.char=""), taxa_are_rows=T)
OTU100 <- otu_table(read.table("data/100_otus.tsv", header=T, check.names=F, row.names=1,
                               skip=1, sep="\t", comment.char=""), taxa_are_rows=T)
# Import taxonomy
TAX97 <- tax_table(as.matrix(read.table("data/97_otus_tax_assignments.txt", sep="\t", row.names=2,
                                        col.names=c("Clade", "OTU.ID", "Taxonomy", "Eval", "Subtype"))))
TAX97bs <- tax_table(as.matrix(read.table("data/97_otus_bysample_tax_assignments.txt", sep="\t", row.names=2,
                                        col.names=c("Clade", "OTU.ID", "Taxonomy", "Eval", "Subtype"))))
TAX97bsp <- tax_table(as.matrix(read.table("data/97_otus_byspecies_tax_assignments.txt", sep="\t", row.names=2,
                                        col.names=c("Clade", "OTU.ID", "Taxonomy", "Eval", "Subtype"))))
TAX100 <- tax_table(as.matrix(read.table("data/100_otus_tax_assignments.txt", sep="\t", row.names=2,
                                         col.names=c("Clade", "OTU.ID", "Taxonomy", "Eval", "Subtype"))))
# Import sample data
SAM <- sample_data(read.delim("data/mapping_file.txt", sep="\t", header=T, row.names=1))
# Build phyloseq objects
phy97 <- phyloseq(OTU97, TAX97, SAM)
phy97bs <- phyloseq(OTU97bs, TAX97bs, SAM)
phy97bsp <- phyloseq(OTU97bsp, TAX97bsp, SAM)
phy100 <- phyloseq(OTU100, TAX100, SAM)
```

### Filter dataset
```{r filter_dataset, cache=T}
# Filter OTUs by minimum count
# Set threshold count
n <- 10
# Identify OTUs below threshold count
taxa97 <- taxa_sums(phy97)[which(taxa_sums(phy97) >= n)]
taxa97bs <- taxa_sums(phy97bs)[which(taxa_sums(phy97bs) >= n)]
taxa97bsp <- taxa_sums(phy97bsp)[which(taxa_sums(phy97bsp) >= n)] 
taxa100 <- taxa_sums(phy100)[which(taxa_sums(phy100) >= n)]
# Remove taxa below threshold count
phy97.f <- prune_taxa(names(taxa97), phy97)
phy97bs.f <- prune_taxa(names(taxa97bs), phy97bs)
phy97bsp.f <- prune_taxa(names(taxa97bsp), phy97bsp)
phy100.f <- prune_taxa(names(taxa100), phy100)

@@

par(mfrow=c(1,1), mar=c(2, 1.5, 2, 5), lwd=0.1, cex=0.7)
# Plot composition of 97% within-sample OTUs colored by clade
composition(phy97bs.f.p, col=taxcolors[factor(data.frame(tax_table(phy97bs.f.p))[order(data.frame(tax_table(phy97bs.f.p))$Subtype), ]$Clade, levels=c("A","B","C","D","F","G"))], legend=T)
# Plot composition of 97% within-sample OTUs colored by OTU
#composition(phy97bs.f.p, col=rainbow(ntaxa(phy97bs.f.p)), legend=F)

@@

## *Diploria strigosa*  
Notice how 97% OTUs and 97% within-sample OTUs are **NOT THE SAME!** Specifically, notice how in the 97% OTU dataset, all samples except two are dominated by the same OTU, denovo235. In the 97% within-sample OTU dataset, those samples are dominated by three different OTUs -- denovo73, denovo231, and denovo10. Look at the 100% OTU dataset to see the composition of unique sequence variants in these samples. Indeed, these samples are dominated by **different sequence variants**. However, in the 97% OTU approach, they have all been collapsed into the same OTU, while in the 97% within-sample clustering approach, samples with different dominant sequence variants are maintained as separate OTUs, even though these different sequence variants are more than 97% similar to each other. I suggest that the  within-sample clustering approach is a better way of treating *Symbiodinium* ITS2 sequence data.
```{r dstr, cache=T, fig.height=16, fig.width=8}
# Create subsetted phyloseq objects for Diploria strigosa
dstr97.f.p <- subset_samples(phy97.f.p, Species=="strigosa")
dstr97bs.f.p <- subset_samples(phy97bs.f.p, Species=="strigosa")
dstr97bsp.f.p <- subset_samples(phy97bsp.f.p, Species=="strigosa")
dstr100.f.p <- subset_samples(phy100.f.p, Species=="strigosa")
# Plot custom barplots for Diploria strigosa
par(mfrow=c(4,1), mar=c(4,4,2,0), mgp=c(0,0.5,0), lwd=0.1)
otubarplot(dstr97.f.p, main="97% OTUs")
otubarplot(dstr97bsp.f.p, main="97% within-species OTUs")
otubarplot(dstr97bs.f.p, main="97% within-sample OTUs")
otubarplot(dstr100.f.p, main="100% OTUs")
```

@@

# Best one for now ---------
V(net)$shape <- c("square", "circle")[factor(V(net)$type)]
V(net)$size <- ifelse(is.na(V(net)$Clade), 15, 10*sqrt(degree(net))) #5*sqrt(degree(net))
V(net)$label <- ifelse(is.na(V(net)$Clade), names(V(net)),  
                       unlist(lapply(strsplit(V(net)$Subtype, split="_"), "[", 1)))
V(net)$label <- str_replace_all(V(net)$label, c("alcicornis"="M.alc.", "annularis"="O.ann",
                                                "astreoides"="P.ast.", "cavernosa"="M.cav.",
                                                "cylindrus"="D.cyl.", "fragum"="F.fra.", 
                                                "furcata"="P.fur.", "radians"="S.rad.", 
                                                "siderea"="S.sid.", "strigosa"="D.str."))
V(net)$color <- ifelse(is.na(V(net)$Clade), "white",
                       taxcolors[factor(V(net)$Clade, levels=c("A","B","C","D","F","G"))])
E(net)$color <- "gray60"
E(net)$width <- 15 * E(net)$weight
set.seed(12347)
set.seed(12364)
set.seed(12374)
l <- layout.fruchterman.reingold(net)
l <- norm_coords(l, ymin=-1, ymax=1, xmin=-1, xmax=1)
par(mar=c(0,0,0,0))
plot(net, rescale=F, layout=l*1, edge.curved=0.1, vertex.label.cex=V(net)$size/20,
     vertex.label.color="black")
#legend(x=-1.05, y=1.1, c("CladeA", "CladeB", "CladeC", "CladeD", "CladeF", "CladeG"), pch=21,
#     col="#777777", pt.bg=taxcolors, pt.cex=1.1, cex=.6, bty="n", ncol=2, text.width=0.2)
```

@@

# Best one for now ---------
V(net)$shape <- c("square", "circle")[factor(V(net)$type)]
V(net)$size <- ifelse(is.na(V(net)$Clade), 15, 10*(degree(net))^0.75) #5*sqrt(degree(net))
V(net)$label <- ifelse(is.na(V(net)$Clade), names(V(net)),  
                       unlist(lapply(strsplit(V(net)$Subtype, split="_"), "[", 1)))
V(net)$label <- str_replace_all(V(net)$label, c("alcicornis"="M.alc.", "annularis"="O.ann",
                                                "astreoides"="P.ast.", "cavernosa"="M.cav.",
                                                "cylindrus"="D.cyl.", "fragum"="F.fra.", 
                                                "furcata"="P.fur.", "radians"="S.rad.", 
                                                "siderea"="S.sid.", "strigosa"="D.str."))
V(net)$color <- ifelse(is.na(V(net)$Clade), "white",
                       taxcolors[factor(V(net)$Clade, levels=c("A","B","C","D","F","G"))])
E(net)$color <- "gray60"
E(net)$width <- 15 * E(net)$weight
set.seed(12347)
set.seed(12364)
set.seed(12374)
l <- layout.fruchterman.reingold(net)
l <- norm_coords(l, ymin=-1, ymax=1, xmin=-1, xmax=1)
par(mar=c(0,0,0,0))
plot(net, rescale=F, layout=l*1, edge.curved=0.1, vertex.label.cex=V(net)$size/20,
     vertex.label.color="black")
#legend(x=-1.05, y=1.1, c("CladeA", "CladeB", "CladeC", "CladeD", "CladeF", "CladeG"), pch=21,
#     col="#777777", pt.bg=taxcolors, pt.cex=1.1, cex=.6, bty="n", ncol=2, text.width=0.2)
```

## Background symbionts
#### (<1% relative abundance)
```{r bk.net, fig.width=5, fig.height=5, fig.align="center"}
# filter out OTUs that only occurred in one sample
phy97bs.f.p2 <- filter_taxa(phy97bs.f.p, function(x) sum(x>0) > 1, prune=TRUE)
# Get network with one node for each species.
otudf <- data.frame(t(otu_table(phy97bs.f.p2)), check.names=F)
# Percent of samples of species that contain OTU at less than 1% relative abundance (but more than 0%)

sp.ag <- aggregate(otudf, by=list(Species=data.frame(sample_data(phy97bs.f.p2))$Species), 
                   FUN=function(x) any(x>0 & x<0.01))
rownames(sp.ag) <- sp.ag$Species
sp.ag <- sp.ag[, -1]
edges <- setNames(melt(as.matrix(sp.ag)), nm=c("source", "target", "weight"))  # Melt data
edges <- droplevels(edges[edges$weight>0, ])
# Create "nodes" table from sample and tax data
sdf <- data.frame(id=levels(data.frame(sample_data(phy97bs.f.p2))$Species))
tdf <- data.frame(tax_table(phy97bs.f.p2))[rownames(data.frame(tax_table(phy97bs.f.p2))) %in% edges$target, ]
tdf$id <- rownames(tdf)
nodes <- merge(sdf, tdf, all=T)
nodes$type <- ifelse(is.na(nodes$Clade), 1, 2)
# Create network
net <- graph_from_data_frame(d=edges, vertices=nodes, directed=F)

# Edit vertex attributes
V(net)$shape <- c("square", "circle")[factor(V(net)$type)]
V(net)$size <- ifelse(is.na(V(net)$Clade), 15, 5*degree(net)) #5*sqrt(degree(net))
V(net)$label <- ifelse(is.na(V(net)$Clade), names(V(net)),  
                       unlist(lapply(strsplit(V(net)$Subtype, split="_"), "[", 1)))
V(net)$label <- str_replace_all(V(net)$label, c("alcicornis"="M.alc.", "annularis"="O.ann",
                                                "astreoides"="P.ast.", "cavernosa"="M.cav.",
                                                "cylindrus"="D.cyl.", "fragum"="F.fra.", 
                                                "furcata"="P.fur.", "radians"="S.rad.", 
                                                "siderea"="S.sid.", "strigosa"="D.str."))
V(net)$color <- ifelse(is.na(V(net)$Clade), "white",
                       taxcolors[factor(V(net)$Clade, levels=c("A","B","C","D","F","G"))])
E(net)$color <- "gray60"
E(net)$width <- E(net)$weight * 2

# Merge D nodes (assuming they are all D1a/trenchii)
# Get original node map (sequence starting at one to the total number of nodes)
nodemap <- seq(1:length(V(net)))
# Make a copy of the nodemap that will be modified
nodemapnew <- nodemap
# Find all nodes beginning with D, and give them the same node ID number
nodemapnew[grep("^D", V(net)$Subtype)] <- min(grep("^D", V(net)$Subtype))
# Replace the now missing values in the node map so node IDs remain sequential
while (max(setdiff(nodemap, nodemapnew)) < max(nodemap)) {
  nodemapnew[which(nodemapnew==max(setdiff(nodemap, nodemapnew))+1)] <- min(setdiff(nodemap, nodemapnew))
}
# Contract the network using the new nodemap (nodes with same node ID are merged)
net2 <- contract(net, nodemapnew, vertex.attr.comb=list(size="sum", "last"))
# Simplify network so that multiple edges are combined
net2 <- simplify(net2, remove.multiple=T, edge.attr.comb=list(weight="mean", width="mean", "ignore"))

@@

## 97% within-sample OTUs
```{r betad97bs, cache=T, fig.width=6, fig.height=6}
betad97bs <- betad(phy97bs.f.t, group="Species")
with(betad97bs$sambdsumm.ord, {
  plot(mean, type="n", main="97% within-sample OTUs",
       ylim=c(0, 0.75), ylab="Distance to centroid", xaxt="n", xlab="")
  arrows(1:10, mean - se, 1:10, mean + se, length=0.05, angle=90, code=3)
  points(1:10, mean, cex=1, pch=21, bg="white")
  text(1:10, par("usr")[3]-0.025, srt=45, adj=1, xpd=T, cex=0.75, 
       labels=levels(SAM$Species)[order(betad97bs$sambdsumm$mean, decreasing=T)])
  text(1:10, mean + se + 0.03, labels=betad97bs$saml$Letters[as.character(group)], cex=0.5)
})
```

## 100% OTUs
```{r betad100, cache=T, fig.width=6, fig.height=6}
betad100 <- betad(phy100.f.t, group="Species")
with(betad100$sambdsumm.ord, {
  plot(mean, type="n", main="100% OTUs",
       ylim=c(0, 0.75), ylab="Distance to centroid", xaxt="n", xlab="")
  arrows(1:10, mean - se, 1:10, mean + se, length=0.05, angle=90, code=3)
  points(1:10, mean, cex=1, pch=21, bg="white")
  text(1:10, par("usr")[3]-0.025, srt=45, adj=1, xpd=T, cex=0.75, 
       labels=levels(SAM$Species)[order(betad100$sambdsumm$mean, decreasing=T)])
  text(1:10, mean + se + 0.03, labels=betad100$saml$Letters[as.character(group)], cex=0.5)
})
```

>>>>

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = normalizePath(".."))
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 3, digits = 9)
```

```{r setup_2, include=FALSE}
# Load all R scripts in "R/" directory
sapply(list.files(path="R", pattern="*[^w].R", full.names=TRUE), source, .GlobalEnv)
# Load package libraries
library(phyloseq); library(vegan); library(multcompView); library(reshape2); library(igraph); library(stringr)
# Set colors for plotting clades
taxcolors <- matrix(c("#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462"), 
                    dimnames=list(c("CladeA", "CladeB", "CladeC", "CladeD", "CladeF", "CladeG")))
```

@@

# Data pre-processing
### Import dataset
The [phyloseq](https://joey711.github.io/phyloseq/) package is used here to combine the OTU table, taxonomic assignments, and sample metadata into a single R data object (class 'phyloseq') to facilitate downstream analyses.
```{r data_import, cache=T}

```{r filter_taxa}
# Import taxonomic assignment data from blast and nw
blast.res.colnames <- c("otu", "BLAST.hit", "pct.id", "B.length", "B.mismatches", "B.gapopen", 
                        "q.start", "q.end", "s.start", "s.end", "e.val", "bit.score")
tax.import <- function(b, nw) {
  m <- merge(read.table(b, col.names=blast.res.colnames, stringsAsFactors=FALSE),
             read.table(nw, stringsAsFactors=FALSE), by="otu")
  rownames(m) <- m$otu
  m$Clade <- str_extract(m$nw.hits, pattern="^.")
  return(m)
}

tax97 <- tax.import(b="data/otus_97/blast_tophits.txt", nw="data/otus_97/nw_tophits.tsv")
tax100 <- tax.import(b="data/otus_100/blast_tophits.txt", nw="data/otus_100/nw_tophits.tsv")
tax97bs <- tax.import(b="data/otus_97_bysample/blast_tophits.txt", nw="data/otus_97_bysample/nw_tophits.tsv")
tax97bsp <- tax.import(b="data/otus_97_byspecies/blast_tophits.txt", nw="data/otus_97_byspecies/nw_tophits.tsv")

# Deal with identical taxonomic assignments (because some reference sequences are not unique...)
# Create names for groups of identical sequences based on members of group...
ident <- readLines("data/ITS2db_trimmed_notuniques_otus.txt")
ident <- gsub("denovo[0-9]*\t", "", ident)
ident <- strsplit(ident, split="\t")
ident2 <- lapply(ident, str_match_all, pattern="[A-I]{1}[0-9]{1,3}.*[_/]")
ident2 <- lapply(ident2, function(g) gsub(pattern="\\..*_$", "", x=unlist(g)))
ident2 <- lapply(ident2, function(g) gsub(pattern="_$", "", g))
ident2 <- lapply(ident2, function(g) unlist(strsplit(g, split="/")))
subtypes <- lapply(ident2, function(x) levels(factor(unlist(x))))
subtypes <- lapply(subtypes, function(s) paste(paste(s, collapse="/"), "_multiple", sep=""))
names(ident) <- subtypes
ident <- melt(do.call(rbind, ident))
ident <- unique(ident[order(ident[,1]), c(3,1)])

# Replace any sequence in taxonomy assignment that is a member of a group of identical sequences with the name of the group
collapse.idents <- function(df) {
  within(df, {
    for (i in 1:nrow(ident)) {
      BLAST.hit <- gsub(ident[i,1], ident[i,2], BLAST.hit)
      nw.hits <- gsub(ident[i,1], ident[i,2], nw.hits)
      nw.hit1 <- gsub(ident[i,1], ident[i,2], nw.hit1)
    }
  })
}

tax97 <- collapse.idents(tax97)
tax97bs <- collapse.idents(tax97bs)
tax97bsp <- collapse.idents(tax97bsp)
tax100 <- collapse.idents(tax100)

# Get subtypes without accession numbers
get.st <- function(df) {
  within(df, {
    bl.st <- gsub(BLAST.hit, pattern="_.*$", replacement="")
    bl.st <- ifelse(nw.sim==100, bl.st, paste("'", bl.st, "'", sep=""))
    nw.st <- gsub(nw.hits, pattern="_[A-Z]{2}[0-9]{6}", replacement="")
    nw.st <- gsub(nw.st, pattern="_multiple", replacement="")
    nw.st <- ifelse(nw.sim==100, nw.st, paste("'", nw.st, "'", sep=""))
    Subtype <- nw.st
  })
}

tax97 <- get.st(tax97)
tax97bs <- get.st(tax97bs)
tax97bsp <- get.st(tax97bsp)
tax100 <- get.st(tax100)
```

```{r build_phyloseq}
# Import OTU tables
otu97 <- otu_table(read.table("data/otus_97/97_otus.tsv", header=T, check.names=F, row.names=1,
                              skip=1, sep="\t", comment.char=""), taxa_are_rows=T)
otu97bs <- otu_table(read.table("data/otus_97_bysample/97_otus_bysample.tsv", header=T, check.names=F, row.names=1,
                              skip=1, sep="\t", comment.char=""), taxa_are_rows=T)
otu97bsp <- otu_table(read.table("data/otus_97_byspecies/97_otus_byspecies.tsv", header=T, check.names=F, row.names=1,
                              skip=1, sep="\t", comment.char=""), taxa_are_rows=T)
otu100 <- otu_table(read.table("data/otus_100/100_otus.tsv", header=T, check.names=F, row.names=1,
                               skip=1, sep="\t", comment.char=""), taxa_are_rows=T)

# Import sample data
sam <- sample_data(read.delim("data/mapping_file.txt", sep="\t", header=T, row.names=1))

# Build phyloseq objects
phy97 <- phyloseq(otu97, tax_table(as.matrix(tax97)), sam)
phy97bs <- phyloseq(otu97bs, tax_table(as.matrix(tax97bs)), sam)
phy97bsp <- phyloseq(otu97bsp, tax_table(as.matrix(tax97bsp)), sam)
phy100 <- phyloseq(otu100, tax_table(as.matrix(tax100)), sam)
```

### Filter dataset
```{r filter_dataset, cache=T}
# FILTER OUT OTUS THAT ARE NOT SYMBIODINIUM
# Build function to BLAST all sequences with less than 90% similarity to Symbiodinium to the NCBI nr database, and get the top hit. If the top hit does not contain the string "Symbiodinium", then this sequence is assumed to not be Symbiodinium.
is.it.symbio <- function(df, repset) {
  poor.otus <- df[which(df$nw.sim < 90), "otu"]  # Identify OTUs with <90% similarity to Symbiodinium
  sapply(poor.otus, function(otu) {  # Get sequences for these OTUs to BLAST to NCBI
    system(paste("awk '/", otu, " /{print;getline;print}' ", repset, " >> temp.fasta", sep=""))
  })
  blast.res <- system("blastn -db nr -remote -query temp.fasta -outfmt '6 qseqid qcovs stitle sseqid' -max_target_seqs 1 | sort -u", intern=TRUE)  # Blast sequences to NCBI
  blast.res <- data.frame(do.call(rbind, lapply(strsplit(blast.res, split="\t"), "[", c(1,3))),
                          stringsAsFactors=FALSE)
  blast.res$sym <- str_detect(blast.res[, 2], "Symbiodinium")  # See if Symbiodinium is in BLAST result
  df2 <- setNames(merge(poor.otus, blast.res, by=1, all=T), nm=c("otu", "hit", "sym"))
  df2[which(is.na(df2$hit)), c("hit", "sym")] <- c("No hit", FALSE)
  system("rm temp.fasta")  # Cleanup temp file
  return(df2)
}

tax97.blasts <- is.it.symbio(tax97, "data/otus_97/97_otus_rep_set.fasta")
tax97.notsym <- as.character(tax97.blasts[which(tax97.blasts$sym==FALSE), "otu"])

tax97bs.blasts <- is.it.symbio(tax97bs, "data/otus_97_bysample/all_rep_set_rep_set.fasta")
tax97bs.notsym <- as.character(tax97bs.blasts[which(tax97bs.blasts$sym==FALSE), "otu"])

tax97bsp.blasts <- is.it.symbio(tax97bsp, "data/otus_97_byspecies/all_rep_set_rep_set.fasta")
tax97bsp.notsym <- as.character(tax97bsp.blasts[which(tax97bsp.blasts$sym==FALSE), "otu"])

tax100.blasts <- is.it.symbio(tax100, "data/otus_100/100_otus_rep_set.fasta")
tax100.notsym <- as.character(tax100.blasts[which(tax100.blasts$sym==FALSE), "otu"])

# Remove OTUs that are not Symbiodinium
phy97.f <- prune_taxa(!taxa_names(phy97) %in% tax97.notsym, phy97)
phy97bs.f <- prune_taxa(!taxa_names(phy97bs) %in% tax97bs.notsym, phy97bs)
phy97bsp.f <- prune_taxa(!taxa_names(phy97bsp) %in% tax97bsp.notsym, phy97bsp)
phy100.f <- prune_taxa(!taxa_names(phy100) %in% tax100.notsym, phy100)

# Filter OTUs by minimum count
# Set threshold count
n <- 10
# Identify OTUs below threshold count
taxa97 <- taxa_sums(phy97.f)[which(taxa_sums(phy97.f) >= n)]
taxa97bs <- taxa_sums(phy97bs.f)[which(taxa_sums(phy97bs.f) >= n)]
taxa97bsp <- taxa_sums(phy97bsp.f)[which(taxa_sums(phy97bsp.f) >= n)] 
taxa100 <- taxa_sums(phy100.f)[which(taxa_sums(phy100.f) >= n)]
# Remove taxa below threshold count
phy97.f <- prune_taxa(names(taxa97), phy97.f)
phy97bs.f <- prune_taxa(names(taxa97bs), phy97bs.f)
phy97bsp.f <- prune_taxa(names(taxa97bsp), phy97bsp.f)
phy100.f <- prune_taxa(names(taxa100), phy100.f)

@@

par(mfrow=c(1,1), mar=c(2, 1.5, 2, 5), lwd=0.1, cex=0.7)
# Plot composition of 97% within-sample OTUs colored by clade
composition(phy97bs.f.p, col=taxcolors[factor(data.frame(tax_table(phy97bs.f.p))[order(data.frame(tax_table(phy97bs.f.p))$nw.st), ]$Clade, levels=c("A","B","C","D","F","G"))], legend=T)
# Plot composition of 97% within-sample OTUs colored by OTU
#composition(phy97bs.f.p, col=rainbow(ntaxa(phy97bs.f.p)), legend=F)

@@

## *Diploria strigosa*  
Notice how 97% OTUs and 97% within-sample OTUs are **NOT THE samE!** Specifically, notice how in the 97% OTU dataset, all samples except two are dominated by the same OTU, denovo235. In the 97% within-sample OTU dataset, those samples are dominated by three different OTUs -- denovo73, denovo231, and denovo10. Look at the 100% OTU dataset to see the composition of unique sequence variants in these samples. Indeed, these samples are dominated by **different sequence variants**. However, in the 97% OTU approach, they have all been collapsed into the same OTU, while in the 97% within-sample clustering approach, samples with different dominant sequence variants are maintained as separate OTUs, even though these different sequence variants are more than 97% similar to each other. I suggest that the  within-sample clustering approach is a better way of treating *Symbiodinium* ITS2 sequence data.
```{r dstr, cache=T, fig.height=16, fig.width=8}
# Create subsetted phyloseq objects for Diploria strigosa
dstr97.f.p <- subset_samples(phy97.f.p, Species=="strigosa")
dstr97bs.f.p <- subset_samples(phy97bs.f.p, Species=="strigosa")
dstr97bsp.f.p <- subset_samples(phy97bsp.f.p, Species=="strigosa")
dstr100.f.p <- subset_samples(phy100.f.p, Species=="strigosa")
# Plot custom barplots for Diploria strigosa
par(mfrow=c(4,1), mar=c(4,4,2,0), mgp=c(0,0.5,0), lwd=0.1)
otubarplot(dstr97.f.p, main="97% OTUs")
otubarplot(dstr97bsp.f.p, main="97% within-species OTUs")
otubarplot(dstr97bs.f.p, main="97% within-sample OTUs")
otubarplot(dstr100.f.p, main="100% OTUs")
```

@@

# Best one for now ---------
V(net)$shape <- c("square", "circle")[factor(V(net)$type)]
V(net)$size <- ifelse(is.na(V(net)$Clade), 15, 10*sqrt(degree(net))) #5*sqrt(degree(net))
V(net)$label <- ifelse(is.na(V(net)$Clade), names(V(net)),  
                       unlist(lapply(strsplit(V(net)$nw.st, split="_"), "[", 1)))
V(net)$label <- str_replace_all(V(net)$label, c("alcicornis"="M.alc.", "annularis"="O.ann",
                                                "astreoides"="P.ast.", "cavernosa"="M.cav.",
                                                "cylindrus"="D.cyl.", "fragum"="F.fra.", 
                                                "furcata"="P.fur.", "radians"="S.rad.", 
                                                "siderea"="S.sid.", "strigosa"="D.str."))
V(net)$color <- ifelse(is.na(V(net)$Clade), "white",
                       taxcolors[factor(V(net)$Clade, levels=c("A","B","C","D","F","G"))])
E(net)$color <- "gray60"
E(net)$width <- 15 * E(net)$weight
set.seed(12347)
set.seed(12364)
set.seed(12374)
l <- layout.fruchterman.reingold(net)
l <- norm_coords(l, ymin=-1, ymax=1, xmin=-1, xmax=1)
par(mar=c(0,0,0,0))
plot(net, rescale=F, layout=l*1, edge.curved=0.1, vertex.label.cex=V(net)$size/20,
     vertex.label.color="black")
#legend(x=-1.05, y=1.1, c("CladeA", "CladeB", "CladeC", "CladeD", "CladeF", "CladeG"), pch=21,
#     col="#777777", pt.bg=taxcolors, pt.cex=1.1, cex=.6, bty="n", ncol=2, text.width=0.2)
```

@@

# Best one for now ---------
V(net)$shape <- c("square", "circle")[factor(V(net)$type)]
V(net)$size <- ifelse(is.na(V(net)$Clade), 15, 10*(degree(net))^0.75) #5*sqrt(degree(net))
V(net)$label <- ifelse(is.na(V(net)$Clade), names(V(net)),  
                       unlist(lapply(strsplit(V(net)$nw.st, split="_"), "[", 1)))
V(net)$label <- str_replace_all(V(net)$label, c("alcicornis"="M.alc.", "annularis"="O.ann",
                                                "astreoides"="P.ast.", "cavernosa"="M.cav.",
                                                "cylindrus"="D.cyl.", "fragum"="F.fra.", 
                                                "furcata"="P.fur.", "radians"="S.rad.", 
                                                "siderea"="S.sid.", "strigosa"="D.str."))
V(net)$color <- ifelse(is.na(V(net)$Clade), "white",
                       taxcolors[factor(V(net)$Clade, levels=c("A","B","C","D","F","G"))])
E(net)$color <- "gray60"
E(net)$width <- 15 * E(net)$weight
set.seed(12347)
set.seed(12364)
set.seed(12374)
l <- layout.fruchterman.reingold(net)
l <- norm_coords(l, ymin=-1, ymax=1, xmin=-1, xmax=1)
par(mar=c(0,0,0,0))
plot(net, rescale=F, layout=l*1, edge.curved=0.1, vertex.label.cex=V(net)$size/20,
     vertex.label.color="black")
#legend(x=-1.05, y=1.1, c("CladeA", "CladeB", "CladeC", "CladeD", "CladeF", "CladeG"), pch=21,
#     col="#777777", pt.bg=taxcolors, pt.cex=1.1, cex=.6, bty="n", ncol=2, text.width=0.2)
```

## Background symbionts
#### (<1% relative abundance)
```{r bk.net, fig.width=5, fig.height=5, fig.align="center"}
# filter out OTUs that only occurred in one sample
phy97bs.f.p2 <- filter_taxa(phy97bs.f.p, function(x) sum(x>0) > 1, prune=TRUE)
# Get network with one node for each species.
otudf <- data.frame(t(otu_table(phy97bs.f.p2)), check.names=F)
# Whether any samples of species contain OTU at less than 1% relative abundance (but more than 0%)
sp.ag <- aggregate(otudf, by=list(Species=data.frame(sample_data(phy97bs.f.p2))$Species), 
                   FUN=function(x) any(x>0 & x<0.01))
rownames(sp.ag) <- sp.ag$Species
sp.ag <- sp.ag[, -1]
edges <- setNames(melt(as.matrix(sp.ag)), nm=c("source", "target", "weight"))  # Melt data
edges <- droplevels(edges[edges$weight>0, ])  # only keep if was observed in more than N colony of a spp
# Create "nodes" table from sample and tax data
sdf <- data.frame(id=levels(data.frame(sample_data(phy97bs.f.p2))$Species))
tdf <- data.frame(tax_table(phy97bs.f.p2))[rownames(data.frame(tax_table(phy97bs.f.p2))) %in% edges$target, ]
tdf$id <- rownames(tdf)
nodes <- merge(sdf, tdf, all=T)
nodes$type <- ifelse(is.na(nodes$Clade), 1, 2)
# Create network
net <- graph_from_data_frame(d=edges, vertices=nodes, directed=F)

# Edit vertex attributes
V(net)$shape <- c("square", "circle")[factor(V(net)$type)]
V(net)$size <- ifelse(is.na(V(net)$Clade), 15, 5*degree(net)) #5*sqrt(degree(net))
V(net)$label <- ifelse(is.na(V(net)$Clade), names(V(net)),  
                       unlist(lapply(strsplit(V(net)$nw.st, split="_"), "[", 1)))
V(net)$label <- str_replace_all(V(net)$label, c("alcicornis"="M.alc.", "annularis"="O.ann",
                                                "astreoides"="P.ast.", "cavernosa"="M.cav.",
                                                "cylindrus"="D.cyl.", "fragum"="F.fra.", 
                                                "furcata"="P.fur.", "radians"="S.rad.", 
                                                "siderea"="S.sid.", "strigosa"="D.str."))
V(net)$color <- ifelse(is.na(V(net)$Clade), "white",
                       taxcolors[factor(V(net)$Clade, levels=c("A","B","C","D","F","G"))])
E(net)$color <- "gray60"
E(net)$width <- E(net)$weight * 2

# Merge D nodes (assuming they are all D1a/trenchii)
# Get original node map (sequence starting at one to the total number of nodes)
nodemap <- seq(1:length(V(net)))
# Make a copy of the nodemap that will be modified
nodemapnew <- nodemap
# Find all nodes beginning with D, and give them the same node ID number
nodemapnew[grep("^D", V(net)$nw.st)] <- min(grep("^D", V(net)$nw.st))
# Replace the now missing values in the node map so node IDs remain sequential
while (max(setdiff(nodemap, nodemapnew)) < max(nodemap)) {
  nodemapnew[which(nodemapnew==max(setdiff(nodemap, nodemapnew))+1)] <- min(setdiff(nodemap, nodemapnew))
}
# Contract the network using the new nodemap (nodes with same node ID are merged)
net2 <- contract(net, nodemapnew, vertex.attr.comb=list(size="sum", "last"))
# Simplify network so that multiple edges are combined
net2 <- simplify(net2, remove.multiple=T, edge.attr.comb=list(weight="mean", width="mean", "ignore"))

@@

## 97% within-sample OTUs
```{r betad97bs, cache=T, fig.width=6, fig.height=6}
betad97bs <- betad(phy97bs.f.t, group="Species")
with(betad97bs$sambdsumm.ord, {
  plot(mean, type="n", main="97% within-sample OTUs",
       ylim=c(0, 0.75), ylab="Distance to centroid", xaxt="n", xlab="")
  arrows(1:10, mean - se, 1:10, mean + se, length=0.05, angle=90, code=3)
  points(1:10, mean, cex=1, pch=21, bg="white")
  text(1:10, par("usr")[3]-0.025, srt=45, adj=1, xpd=T, cex=0.75, 
       labels=levels(sam$Species)[order(betad97bs$sambdsumm$mean, decreasing=T)])
  text(1:10, mean + se + 0.03, labels=betad97bs$saml$Letters[as.character(group)], cex=0.5)
})
```

## 100% OTUs
```{r betad100, cache=T, fig.width=6, fig.height=6}
betad100 <- betad(phy100.f.t, group="Species")
with(betad100$sambdsumm.ord, {
  plot(mean, type="n", main="100% OTUs",
       ylim=c(0, 0.75), ylab="Distance to centroid", xaxt="n", xlab="")
  arrows(1:10, mean - se, 1:10, mean + se, length=0.05, angle=90, code=3)
  points(1:10, mean, cex=1, pch=21, bg="white")
  text(1:10, par("usr")[3]-0.025, srt=45, adj=1, xpd=T, cex=0.75, 
       labels=levels(sam$Species)[order(betad100$sambdsumm$mean, decreasing=T)])
  text(1:10, mean + se + 0.03, labels=betad100$saml$Letters[as.character(group)], cex=0.5)
})
```
